sWGS validation/protocol tests

### 2019-03-13 Validation Test Report.Rmd
updating

---
title: "sWGS validation/protocol tests"
author: "Sarah Killcoyne"
date: "14 Jan 2019"
output: 
  html_document: 
    toc: yes

@@

dirs = list(
  training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
  val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
  protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/'
)

@@

failed.tbl = failed %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   sample = text_spec(sample, link=paste0('file:///',get.png(sample, paste(plot.dir, pid, sep='/') ))), 
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>% dplyr::select(pid,sample,varMAD_median)

@@

val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename),val, by=c('Samplename'='sample'))

@@

pass.fail = bind_rows(pass.fail, 
  val %>% summarise( Pass= length(which(varMAD_median<=0.011))/nrow(val), Fail=length(which(varMAD_median>0.011))/nrow(val), group='Validation'))
```

@@

speedvac.tbl = speedvac %>% arrange(Speedvac, `Ellie’s Cohort`, `Manual QC`) %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   #Speedvac = cell_spec(Speedvac, "html", background = ifelse(Speedvac == 'Yes', 'lightblue', 'yellow')),
   ID = text_spec(paste(`SLX-ID`, `Index Sequence`,sep='.'),link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   Sample = text_spec(Sample, link=paste0('file:///', get.png(paste0(Sample, '_segmentedCoverage.png'), paste(proc.dir,`Patient ID`,sep='/')))),
   #text_spec(Sample, link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'Pass', "green", "red"), color='white'),
   `Ellie’s Cohort` = cell_spec(`Ellie’s Cohort`, "html", color = ifelse(`Ellie’s Cohort` == 'Yes', "blue", "brown")),
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
  dplyr::select(ID,Sample, Speedvac, `Ellie’s Cohort`, `Manual QC`, varMAD_median) %>% rename(`Ellie’s Cohort` = 'Previously Extracted')

@@

kit.tbl = kit %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(Sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png(sub('-','_',`Index Sequence`),plot.dir))),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= 0.011, "green", "red"))) %>%
  dplyr::select(`Index Sequence`, Sample, Kit, Type, `Manual QC`, varMAD_median)

@@

@@

Currently we are guessing that there is a step during extraction that is causing the problem.  All of Ellie’s samples used in the Speedvac test were extracted prior to use by us, though not necessarily all by Ellie. As the protocol being used was from Ellie's thesis and conversations with her, rather than direct training, it seems possible that something was undocumented/unclear and therefore lost in the mix.

>>>>

---
title: "sWGS validation/protocol tests"
author: "Sarah Killcoyne"
date: "27 Feb 2019"
output: 
  html_document: 
    toc: yes

@@

dirs = list(
  training = '~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_fitted_perPatient',
  val_1 = '~/Data/BarrettsProgressionRisk/Analysis/val_test/',
  protocol = '~/Data/BarrettsProgressionRisk/Analysis/protocol_test/',
  kit = '~/Data/BarrettsProgressionRisk/Analysis/kit_test/',
  ci_protocol = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol/'
)

qc.cutoff = 0.011
```

@@

failed.tbl = failed %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   sample = text_spec(sample, link=paste0('file:///',get.png(sample, paste(plot.dir, pid, sep='/') ))), 
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>% dplyr::select(pid,sample,varMAD_median)

@@

val = right_join(readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/ValidationCohort.xlsx') %>% mutate(Samplename = paste0('SLX-',`SLX-ID`, '.', sub('-','_',`Index Sequence`))) %>% select(`Hospital Research ID`,`Block ID`, `Patient ID`, Samplename, `Sample Type`, Collection),val, by=c('Samplename'='sample'))

@@

pass.fail = bind_rows(pass.fail, 
  val %>% summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(val), Fail=length(which(varMAD_median>qc.cutoff))/nrow(val), group='Validation'))

val %>% group_by(`Sample Type`, Collection ) %>% dplyr::summarise(
  Pass=length(which(varMAD_median<=qc.cutoff)), Fail=length(which(varMAD_median>qc.cutoff))
) %>% kable(caption='Pass/Fail by sample type') %>% kable_styling('bordered', full_width = F)
```

@@

speedvac.tbl = speedvac %>% arrange(Speedvac, `Ellie’s Cohort`, `Manual QC`) %>% rowwise() %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>%
 dplyr::mutate(
   #Speedvac = cell_spec(Speedvac, "html", background = ifelse(Speedvac == 'Yes', 'lightblue', 'yellow')),
   ID = text_spec(paste(`SLX-ID`, `Index Sequence`,sep='.'),link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   Sample = text_spec(Sample, link=paste0('file:///', get.png(paste0(Sample, '_segmentedCoverage.png'), paste(proc.dir,`Patient ID`,sep='/')))),
   #text_spec(Sample, link=paste0('file:///',get.png( paste(`SLX-ID`, sub('-','_',`Index Sequence`),sep='.'), plot.dir ))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'Pass', "green", "red"), color='white'),
   `Ellie’s Cohort` = cell_spec(`Ellie’s Cohort`, "html", color = ifelse(`Ellie’s Cohort` == 'Yes', "blue", "brown")),
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
  dplyr::select(ID,Sample, Speedvac, `Ellie’s Cohort`, `Manual QC`, varMAD_median) %>% rename(`Ellie’s Cohort` = 'Previously Extracted')

@@

kit.tbl = kit %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(Sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png(sub('-','_',`Index Sequence`),plot.dir))),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
  dplyr::select(`Index Sequence`, Sample, Kit, Type, `Manual QC`, varMAD_median)

@@

## CI Protocol

At this point we were guessing that the extraction method itself may be at issue, and that there could be small protocol errors that we were unaware of.  We selected a set of samples that had been previously sequenced by Ellie and Cassandra, and then a number of new samples as well.

Cassandra worked with Anna Piskorz from the Brenton lab at the CI to do the extractions using their protocol and their kits, and Anna then did the library prep as well.  However, again a very high percentage failed.  Most concerning is that the samples from Ellie's cohort failed and the orginal ones done in 2017 by Ellie all looked fine.

```{r}
ci = do.call(bind_rows, lapply(list.files(dirs$ci_protocol, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types='cdddddl')
})) %>% select(sample, varMAD_median) %>% dplyr::mutate( Pass = varMAD_median<=qc.cutoff)

info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/ci_protocol/ci_protocol_trial.xlsx') %>%  
  dplyr::mutate( sample = paste0(`SLX-ID`, '.', sub('-', '_', gsub('tp','', `Index Sequence`))))

ci = left_join(ci, info, by='sample') %>% arrange(Pass)

pass.fail = bind_rows(pass.fail, 
  ci %>% dplyr::summarise( Pass= length(which(varMAD_median<=qc.cutoff))/nrow(ci), Fail=length(which(varMAD_median>qc.cutoff))/nrow(ci), group='CI Protocols'))

plot.dir= '~/Data/BarrettsProgressionRisk/QDNAseq/ci_protocol/SLX-15634'
proc.dir = '~/Data/BarrettsProgressionRisk/Analysis/ci_protocol'

ci.tbl = ci %>% dplyr::mutate(varMAD_median = signif(varMAD_median,4)) %>% rowwise() %>%
 dplyr::mutate(
   Sample = text_spec(sample, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),
   `Manual QC` = cell_spec(`Manual QC`, "html", background = ifelse(`Manual QC` == 'PASS', "green", "red"), color='white'),
   `Computational QC` = cell_spec(Pass, "html", background = ifelse(Pass == T, "green", "red"), color='white'),
   `Index Sequence` = text_spec(`Index Sequence`, link=paste0('file:///',get.png('png',paste(proc.dir,as.character(Number),'plots',sep='/')))),   
   varMAD_median  = cell_spec(varMAD_median, "html", color = ifelse(varMAD_median <= qc.cutoff, "green", "red"))) %>%
  dplyr::select(`Patient ID`,`Index Sequence`, Cohort, `Manual QC`, `Computational QC`, varMAD_median) %>% arrange(Cohort)

rowsE = grep('Ellie',ci.tbl$Cohort)
rowsC = grep('Cass',ci.tbl$Cohort)
rowsN = grep('New', ci.tbl$Cohort)

ci.tbl %>%
  kable(format = "html", escape = F) %>% kable_styling("bordered", full_width = F) %>%
  group_rows("Cohort: Ellie's", rowsE[1], rowsE[length(rowsE)],label_row_css = "background-color: #666; color: #fff;") %>%
  group_rows("Cohort: Cassandra's", rowsC[1], rowsC[length(rowsC)],label_row_css = "background-color: #666; color: #fff;")  %>%
  group_rows("Cohort: New Samples", rowsN[1], rowsN[length(rowsN)],label_row_css = "background-color: #666; color: #fff;")

@@

???
